---
title: "Vessel Tracks + Whale Sightings"
author: "WorthWhale Team"
date: "10/2/2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

&nbsp;
  
**Goal: Match up the photo timestamps from whale sightings data with research vessel GPS coordinates in order to assign coordinates to the whale sightings.**  

&nbsp;

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}

# Load packages
library(tidyverse)
library(lubridate)
library(DBI) # DBI stands for "database interface" and creates a connection between R and other database management systems

```

This is the test code to use for the different whale clusters/vessel tracks:

```{r test load data}

# # Read in vessel track data for April 20, 2017
# vesseltracks <- read.csv("data/GPXtoCSV04-20.csv")
# 
# # Read in whale sightings data for April 20, 2017
# whaleclusters <- read.csv("data/2017_clusters_04-20.csv")
# # if we do a read_csv, we can specify how the indiv columns are translated
#   # look in Hadley book for how this works
#   # when googling, include "tidyverse" not even "R"

```

This is the example code to use for each year:

```{r test code}

# vesseltracks$GPStime <- parse_date_time(vesseltracks$time, "%Y/%m/%d %H:%M:%S", tz = "America/Dominica")
# # Match the date you have in your data, R will spit out the correct date format. For Dominica, you have to set the time zone, otherwise R assumes UTC format
# 
# whaleclusters$GPStime <- parse_date_time(whaleclusters$DateTime, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")
#     # Change the "DateTime" column name to "GPStime" in whaleclusters_0420_2017
# 
# # Create function to get whale cluster time to equal the GPS time in order to connect vessel track lat long to whale cluster lat long
# # Subtract the whale cluster gps time and the vessel tracks gps time 
# # find the absolute value 
# # take the min value of the absolute value
# #
# # apply the fuction to all rows in the whalecluster df
# closest <- sapply(seq(nrow(whaleclusters)), function(i) {which.min(abs(whaleclusters$GPStime[i] - vesseltracks$GPStime))})
# 
# whaleclusters$Lat <- vesseltracks$lat[closest]
# 
# whaleclusters$Long <- vesseltracks$lon[closest]

```

Combine whale sightings data with vessel track location data from 2005-2018:

```{r load 2005-2015 clusters}

# Read in 2005-2015 whale clusters data
whaleclusters0515 <- read.csv("Whale_Clusters/2005_2015_cluster.csv")

# Combine the True Time and Date columns because the DateTime column doesn't have all of the times included, or seconds are missing
whaleclusters0515$TrueDateTime <- paste(whaleclusters0515$Date, "", whaleclusters0515$TrueTime)

nrow(whaleclusters0515) #3559

```

```{r 2005}

# Read in vessel tracks data for 2005
vesseltrack_2005 <- read.csv("Vessel_Tracks/2005_vesseltracks.csv")

# Filter out 2005 from 2005-2015 whale clusters data
whaleclusters2005 <- whaleclusters0515 %>%
  filter(Year == "2005")
nrow(whaleclusters2005) #667

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltrack_2005$GPStime <- parse_date_time(vesseltrack_2005$Time, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2005$GPStime <- parse_date_time(whaleclusters2005$TrueDateTime, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")
    
# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2005 <- sapply(seq(nrow(whaleclusters2005)), function(i) {which.min(abs(whaleclusters2005$GPStime[i] - vesseltrack_2005$GPStime))})

whaleclusters2005$Lat <- vesseltrack_2005$Lat[closest2005]

whaleclusters2005$Long <- vesseltrack_2005$Long[closest2005]

# Select the columns we need
finalwc2005 <- whaleclusters2005 %>%
  select(Cluster, GPStime, Lat, Long)
nrow(finalwc2005) #667

```

```{r 2007}

# Read in vessel tracks data for 2006
vesseltrack_2007 <- read.csv("Vessel_Tracks/2007_vesseltracks.csv")

# Filter out 2007 from 2005-2015 whale clusters data
whaleclusters2007 <- whaleclusters0515 %>%
  filter(Year == "2007")
nrow(whaleclusters2007) #102

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltrack_2007$GPStime <- parse_date_time(vesseltrack_2007$time, "%Y/%m/%d %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2007$GPStime <- parse_date_time(whaleclusters2007$TrueDateTime, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")
 
# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2007 <- sapply(seq(nrow(whaleclusters2007)), function(i) {which.min(abs(whaleclusters2007$GPStime[i] - vesseltrack_2007$GPStime))})

whaleclusters2007$Lat <- vesseltrack_2007$lat[closest2007]

whaleclusters2007$Long <- vesseltrack_2007$lon[closest2007]

# Select the columns we need
finalwc2007 <- whaleclusters2007 %>%
  select(Cluster, GPStime, Lat, Long)
nrow(finalwc2007) #102

```

```{r 2008}

# Read in vessel tracks data for 2008
vesseltrack_2008 <- read.csv("Vessel_Tracks/2008_vesseltracks.csv")

# Filter out 2008 from 2005-2015 whale clusters data
whaleclusters2008 <- whaleclusters0515 %>%
  filter(Year == "2008")
nrow(whaleclusters2008) #544

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltrack_2008$GPStime <- parse_date_time(vesseltrack_2008$time, "%Y/%m/%d %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2008$GPStime <- parse_date_time(whaleclusters2008$TrueDateTime, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")

# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2008 <- sapply(seq(nrow(whaleclusters2008)), function(i) {which.min(abs(whaleclusters2008$GPStime[i] - vesseltrack_2008$GPStime))})

whaleclusters2008$Lat <- vesseltrack_2008$lat[closest2008]

whaleclusters2008$Long <- vesseltrack_2008$lon[closest2008]

# Select the columns we need
finalwc2008 <- whaleclusters2008 %>%
  select(Cluster, GPStime, Lat, Long)
nrow(finalwc2008) #544

```

```{r 2009}

# Read in vessel tracks data for 2009
vesseltrack_2009 <- read.csv("Vessel_Tracks/2009_vesseltracks.csv")

# Filter out 2009 from 2005-2015 whale clusters data
whaleclusters2009 <- whaleclusters0515 %>%
  filter(Year == "2009")
nrow(whaleclusters2009) #213

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltrack_2009$GPStime <- parse_date_time(vesseltrack_2009$time, "%Y/%m/%d %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2009$GPStime <- parse_date_time(whaleclusters2009$TrueDateTime, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")

# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2009 <- sapply(seq(nrow(whaleclusters2009)), function(i) {which.min(abs(whaleclusters2009$GPStime[i] - vesseltrack_2009$GPStime))})

whaleclusters2009$Lat <- vesseltrack_2009$lat[closest2009]

whaleclusters2009$Long <- vesseltrack_2009$lon[closest2009]

# Select the columns we need
finalwc2009 <- whaleclusters2009 %>%
  select(Cluster, GPStime, Lat, Long)
nrow(finalwc2009) #213

```

```{r 2010}

# Read in vessel tracks data for 2010
vesseltrack_2010 <- read.csv("Vessel_Tracks/2010_vesseltracks.csv")

# Filter out 2010 from 2005-2015 whale clusters data
whaleclusters2010 <- whaleclusters0515 %>%
  filter(Year == "2010")
nrow(whaleclusters2010) #970

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltrack_2010$GPStime <- parse_date_time(vesseltrack_2010$time, "%Y/%m/%d %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2010$GPStime <- parse_date_time(whaleclusters2010$TrueDateTime, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")

# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2010 <- sapply(seq(nrow(whaleclusters2010)), function(i) {which.min(abs(whaleclusters2010$GPStime[i] - vesseltrack_2010$GPStime))})

whaleclusters2010$Lat <- vesseltrack_2010$lat[closest2010]

whaleclusters2010$Long <- vesseltrack_2010$lon[closest2010]

# Select the columns we need
finalwc2010 <- whaleclusters2010 %>%
  select(Cluster, GPStime, Lat, Long)
nrow(finalwc2010) #970

```

```{r 2011}

# Read in vessel tracks data for 2011
vesseltrack_2011 <- read.csv("Vessel_Tracks/2011_vesseltracks.csv")

# Filter out 2011 from 2005-2015 whale clusters data
whaleclusters2011 <- whaleclusters0515 %>%
  filter(Year == "2011")
nrow(whaleclusters2011) #317

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltrack_2011$GPStime <- parse_date_time(vesseltrack_2011$time, "%Y/%m/%d %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2011$GPStime <- parse_date_time(whaleclusters2011$TrueDateTime, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")

# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2011 <- sapply(seq(nrow(whaleclusters2011)), function(i) {which.min(abs(whaleclusters2011$GPStime[i] - vesseltrack_2011$GPStime))})

whaleclusters2011$Lat <- vesseltrack_2011$lat[closest2011]

whaleclusters2011$Long <- vesseltrack_2011$lon[closest2011]

# Select the columns we need
finalwc2011 <- whaleclusters2011 %>%
  select(Cluster, GPStime, Lat, Long)
nrow(finalwc2011) #317

```


```{r 2012}

# Read in vessel tracks data for 2012
vesseltrack_2012 <- read.csv("Vessel_Tracks/2012_vesseltracks.csv")

# Filter out 2012 from 2005-2015 whale clusters data
whaleclusters2012 <- whaleclusters0515 %>%
  filter(Year == "2012")
nrow(whaleclusters2012) #173

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltrack_2012$GPStime <- parse_date_time(vesseltrack_2012$time, "%Y/%m/%d %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2012$GPStime <- parse_date_time(whaleclusters2012$TrueDateTime, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")

# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2012 <- sapply(seq(nrow(whaleclusters2012)), function(i) {which.min(abs(whaleclusters2009$GPStime[i] - vesseltrack_2012$GPStime))})

whaleclusters2012$Lat <- vesseltrack_2012$lat[closest2012]

whaleclusters2012$Long <- vesseltrack_2012$lon[closest2012]

# Select the columns we need
finalwc2012 <- whaleclusters2012 %>%
  select(Cluster, GPStime, Lat, Long)
nrow(finalwc2012) #173

```

```{r 2014}

# Read in vessel tracks data for 2014
vesseltrack_2014 <- read.csv("Vessel_Tracks/2014_vesseltracks.csv")

# Filter out 2014 from 2005-2015 whale clusters data
whaleclusters2014 <- whaleclusters0515 %>%
  filter(Year == "2014")
nrow(whaleclusters2014) #81

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltrack_2014$GPStime <- parse_date_time(vesseltrack_2014$time, "%Y/%m/%d %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2014$GPStime <- parse_date_time(whaleclusters2014$TrueDateTime, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")

# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2014 <- sapply(seq(nrow(whaleclusters2014)), function(i) {which.min(abs(whaleclusters2014$GPStime[i] - vesseltrack_2014$GPStime))})

whaleclusters2014$Lat <- vesseltrack_2014$lat[closest2014]

whaleclusters2014$Long <- vesseltrack_2014$lon[closest2014]

# Select the columns we need
finalwc2014 <- whaleclusters2014 %>%
  select(Cluster, GPStime, Lat, Long)
nrow(finalwc2014) #81

```

```{r 2015}

# Read in vessel tracks data for 2015
vesseltrack_2015 <- read.csv("Vessel_Tracks/2015_vesseltracks.csv")

# Filter out 2015 from 2005-2015 whale clusters data
whaleclusters2015 <- whaleclusters0515 %>%
  filter(Year == "2015")
nrow(whaleclusters2015) #492

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltrack_2015$GPStime <- parse_date_time(vesseltrack_2015$time, "%Y/%m/%d %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2015$GPStime <- parse_date_time(whaleclusters2015$TrueDateTime, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")

# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2015 <- sapply(seq(nrow(whaleclusters2015)), function(i) {which.min(abs(whaleclusters2015$GPStime[i] - vesseltrack_2015$GPStime))})

whaleclusters2015$Lat <- vesseltrack_2015$lat[closest2015]

whaleclusters2015$Long <- vesseltrack_2015$lon[closest2015]

# Select the columns we need
finalwc2015 <- whaleclusters2015 %>%
  select(Cluster, GPStime, Lat, Long)
nrow(finalwc2015) #492

```

```{r 2016}

# Read in vessel tracks data for 2016
vesseltracks2016 <- read.csv("Vessel_Tracks/2016_vesseltracks.csv")

# Read in whale clusters data for 2016
whaleclusters2016 <- read.csv("Whale_Clusters/2016_cluster.csv")
nrow(whaleclusters2016) #412

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltracks2016$GPStime <- parse_date_time(vesseltracks2016$time, "%Y/%m/%d %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2016$GPStime <- parse_date_time(whaleclusters2016$DateTime, "%d/%m/%y %H:%M:%S", tz = "America/Dominica")

# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2016 <- sapply(seq(nrow(whaleclusters2016)), function(i) {which.min(abs(whaleclusters2016$GPStime[i] - vesseltracks2016$GPStime))})

whaleclusters2016$Lat <- vesseltracks2016$lat[closest2016]

whaleclusters2016$Long <- vesseltracks2016$lon[closest2016]

# Select the columns we need
finalwc2016 <- whaleclusters2016 %>%
  select(Cluster, GPStime, Lat, Long)
nrow(finalwc2016) #412

```

```{r 2017}

# Read in vessel tracks data for 2017
vesseltracks2017 <- read.csv("Vessel_Tracks/2017_vesseltracks.csv")

# Read in whale clusters data for 2017
whaleclusters2017 <- read.csv("Whale_Clusters/2017_cluster.csv")
nrow(whaleclusters2017) #355

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltracks2017$GPStime <- parse_date_time(vesseltracks2017$time, "%Y/%m/%d %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2017$GPStime <- parse_date_time(whaleclusters2017$DateTime, "%d/%m/%y %H:%M:%S", tz = "America/Dominica")

# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2017 <- sapply(seq(nrow(whaleclusters2017)), function(i) {which.min(abs(whaleclusters2017$GPStime[i] - vesseltracks2017$GPStime))})

whaleclusters2017$Lat <- vesseltracks2017$lat[closest2017]

whaleclusters2017$Long <- vesseltracks2017$lon[closest2017]

# Select the columns we need
finalwc2017 <- whaleclusters2017 %>%
  select(Cluster, GPStime, Lat, Long)
nrow(finalwc2017) #355

```

```{r 2018, warning=FALSE}

# Read in vessel tracks data for 2018
vesseltracks2018 <- read.csv("Vessel_Tracks/2018_vesseltracks.csv")

# Read in whale clusters data for 2018
whaleclusters2018 <- read.csv("Whale_Clusters/2018_cluster.csv")
nrow(whaleclusters2018) #567

# Change the "time" column in original vessel track data to "GPStime" that contains the correct UTC time format. The parse_date_time function converts whatever format your data's date is in currently (%Y=4-digit year, %y=2-digit year, %m=month (00-12), %b=abbreviated month, %B=unabbreviated month, %d=day as a number (0-31), %H=hours, %M=minutes, %S=seconds) into the correct format (UTC). Need to set the time zone for Dominica, otherwise R assumes UTC format.
vesseltracks2018$GPStime <- parse_date_time(vesseltracks2018$time, "%b/%d/%Y %H:%M:%S", tz = "America/Dominica")

# Change the "Datetime" column name in original whale clusters data to "GPStime" that contains the correct UTC time format.
whaleclusters2018$TrueDateTime <- paste(whaleclusters2018$Date, "", whaleclusters2018$TrueTime)
whaleclusters2018$GPStime <- parse_date_time(whaleclusters2018$TrueDateTime, "%d/%m/%Y %H:%M:%S", tz = "America/Dominica")

# Create function to match up the timestamp of sightings in the whale cluster data with the closest GPS time from the vessel track data in order to assign geographic coordinates to the whale cluster data.
# Subtract the whale cluster GPStime and the vessel tracks GPStime.
# Find the absolute value 
# Take the min value of the absolute value
# Apply the fuction to all rows in the whale cluster dataframe
closest2018 <- sapply(seq(nrow(whaleclusters2018)), function(i) {which.min(abs(whaleclusters2018$GPStime[i] - vesseltracks2018$GPStime))})

whaleclusters2018$Lat <- vesseltracks2018$lat[closest2018]

whaleclusters2018$Long <- vesseltracks2018$lon[closest2018]

# Select the columns we need
finalwc2018 <- whaleclusters2018 %>%
  dplyr::select(Cluster, GPStime, Lat, Long)
nrow(finalwc2018) #567

```


Combine all new whale clusters data frames into one data frame/csv that now contains the time stamp and location data for all whale sightings from 2005-2018:

```{r Final whale clusters data 2005-2018}

# Combine all years (2005-2018) into one data frame
whaleclusters0518 <- do.call("rbind", list(finalwc2005, finalwc2007, finalwc2008, finalwc2009, finalwc2010, finalwc2011, finalwc2012, finalwc2014, finalwc2015, finalwc2016, finalwc2017, finalwc2018))

# Write csv
write.csv(whaleclusters0518, "whaleclusters2005_2018.csv")


```

Transfer "whaleclusters2005_2018.csv" to SQLite database:

```{r SQLite database}

whalesightings2005_2018 <- read_csv("whaleclusters2005_2018.csv")

filename <- "db.sqlite"
mydb <- dbConnect(RSQLite::SQLite(), filename)
dbWriteTable(mydb, "whalesightings2005_2018", whalesightings2005_2018)

```

